---
title: "Getting Started with regdatasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with regdatasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4,
  message = FALSE,
  warning = FALSE
)
```

## Installation

Install the development version of **regdatasets** from GitHub:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("joonho112/regdatasets")
```

## Loading Datasets

All 25 datasets are available via the standard `data()` function:

```{r}
library(regdatasets)

# Load a single dataset
data(gcse)
head(gcse)
```

Every dataset is stored as a **tibble**, so it prints cleanly and works
seamlessly with the tidyverse:

```{r}
class(gcse)
```

## Exploring a Dataset

Use `?dataset_name` to view the full documentation, including variable
descriptions, sources, and example analyses:

```{r, eval = FALSE}
?gcse
?berkeley
?crime
```

## Example: Simple Linear Regression

The `gcse` dataset is the primary running example for Chapters 1–2.
It contains GCSE exam scores and London Reading Test (LRT) scores for
4,059 students in 65 London secondary schools:

```{r}
# Fit a simple linear regression
model <- lm(gcse ~ lrt, data = gcse)
summary(model)
```

```{r fig-gcse-scatter, fig.cap = "GCSE scores vs. LRT scores with fitted regression line."}
library(ggplot2)

ggplot(gcse, aes(x = lrt, y = gcse)) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", color = "#E74C3C", se = TRUE) +
  labs(
    title = "GCSE vs. London Reading Test",
    x = "London Reading Test Score (Age 11)",
    y = "GCSE Score (Age 16)"
  ) +
  theme_minimal(base_size = 13)
```

## Example: Logistic Regression

The `berkeley` dataset demonstrates Simpson's paradox — the reversal of
an apparent gender bias after controlling for department:

```{r}
data(berkeley)

# Unadjusted model
crude <- glm(admitted ~ female, family = binomial, data = berkeley)

# Adjusted model (controlling for department)
adjusted <- glm(admitted ~ female + factor(departme),
                family = binomial, data = berkeley)

# Compare odds ratios
exp(coef(crude)["female"])    # Crude OR
exp(coef(adjusted)["female"]) # Adjusted OR
```

The crude odds ratio suggests female applicants are disadvantaged, but
after adjusting for department the effect essentially disappears —
a classic example of confounding.

## Example: One-way ANOVA

The `reading` dataset compares three reading instruction methods:

```{r}
data(reading)

# ANOVA via regression with dummy variables
model_anova <- lm(post1 ~ factor(group), data = reading)
anova(model_anova)
```

```{r fig-reading-box, fig.cap = "Post-test scores by instruction group."}
ggplot(reading, aes(x = factor(group), y = post1, fill = factor(group))) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  scale_fill_manual(values = c("#2C3E50", "#E74C3C", "#3498DB")) +
  labs(
    title = "Reading Comprehension by Instruction Method",
    x = "Instruction Group",
    y = "Post-test Score (post1)"
  ) +
  theme_minimal(base_size = 13)
```

## Available Datasets

For a complete list of all 25 datasets and their chapter mappings, see the
[Dataset Overview](dataset-overview.html) article.

```{r}
# List all datasets in the package
data(package = "regdatasets")$results[, c("Item", "Title")]
```
